---
title: "Portfolio 2: TComm AI Graph"
---

> The goal of this project is to create a graph for my AI research. Specifically, the graph will display our results for our QCV Truthful Communication Scale Validation project. The graph should replicate a line graph created by Mike that has predictions (correlations) along the y-axis and the criterion variables on the x-axis. I will use this graph in my poster at an upcoming conference! 

```{r installhaven}
#install.packages("haven")
```
The haven package allows you to read in SPSS data

```{r loadhaven}
library(haven)
library(tidyr)
library(tidyverse)
```


```{r reading data}
AI <- read_spss("C:/Users/crawhk23/Documents/GitHub/Portfolio/TComm AI & Human SS (Names and No Names).sav")
```

```{r}
print(AI)
```

```{r creatingmeans}
AIAvgs <- AI %>%  
  mutate(
    ExpertAvg = rowMeans(across(mf:es), na.rm = TRUE),
    AIAvg= rowMeans(across(AI1:AI6), na.rm = TRUE)
    )

```

```{r}
print(AIAvgs)
```

```{r plotting}
ggplot(AIAvgs, aes(x = Scale)) +
  geom_line(aes(y = Actual, group = 1, linetype = "Actual Scores")) +
  geom_line(aes(y = ExpertAvg, group = 2, linetype = "Expert Rater Predictions")) +
  geom_line(aes(y = AIAvg, group = 3, linetype = "AI Predictions")) +
  labs(title = "Truthful Communication Scale Validation",
       x = "Criterion Scale", linetype = "Rater", y= "Correlations")+ 
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(),
    axis.ticks = element_line(),
    axis.text.x = element_text(angle = 45, hjust = 1)) + 
scale_y_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.2))


```

