---
title: "Portfolio 3 - Fueling My Titanic Obsession"
---

> The goal of this project is to go through the Titanic lab that I found online and provide commentary on its usefullness. For next portfolio piece I hope to create a new lab using the Titanic data set based on my review of the current project.  https://medium.com/@datathon/r-for-data-science-data-wrangling-d125285a6876 

### Eliminating early "fluff"
> Much of the current lab has information that is unnecessary for where students are at this point in the semester. For example, the site includes installing R directions and explains the layout of R. For a visualization lab for this class, these instructions are unnessecary.

### Importing Data

> The clarity of this step could be improved. Here are some new steps:

1) Go to Kaggle Titanic Dataset - there are a lot of datasets on titanic out there, however the lab was not clear which was being used... so i believe this is the best match 
https://www.kaggle.com/datasets/ibooth1/titanic3 

2) Download the dataset in the format you would like, I did excel csv

3) Extract the data set to your portfolio folder

4) Install/load packages 

```{r}
library(tidyr)
library(readxl)
library(Rcpp) # needed to load Amelia
library(Amelia) # package for missing data/imputation
library(dplyr)
```

5) Read in data 
```{r}
library(readxl)
titanic <- read_xls("titanic3.xls")

```

### Exploring and cleaning dataset 

```{r}
str(titanic)
```
```{r}
as_tibble(titanic)
```


#### Checking for missing values 
```{r}
missmap(titanic, col = c("red", "green"))
```
21% of data is missing 

```{r}
#Select relevant columns for the analysis
selected_titanic <- titanic %>%
  select(age, pclass, sex, survived, embarked, fare, cabin, parch, sibsp, home.dest)
```

Reassess missing map
```{r}
missmap(selected_titanic, col = c("red", "green"))
```
Missing is now 11% 

```{r}
#Drop all the missing values
selected_titanic <- drop_na(selected_titanic)
```
```{r}
missmap(selected_titanic, col = c("red", "green"))
```

```{r}
as_tibble(selected_titanic)
```


```{r}
#merge columns parch and sibsp to create a new column, FamilySize
selected_titanic$FamilySize <- selected_titanic$sibsp
 + selected_titanic$parch + 1

#Categorize the fare column and assign label to each category
selected_titanic$FareCategory <- cut(selected_titanic$fare, breaks = c(0, 10, 20, 50, 100, Inf), labels = c("Lowest", "Lower Middle", "Upper Middle", "Higher", "Highest"))
```

```{r}
View(selected_titanic)
```

```{r}
#Remove irrelevant columns
selected_titanic <- selected_titanic %>% 
  select(-fare, -parch, -sibsp)
```

```{r}
#Change the values of columns survived, age, and pclass

selected_titanic <- selected_titanic %>%
  mutate(
    survived = ifelse(survived == 0, "No", "Yes"),

    age = ifelse(age >= 18, "Adult", "Child"),

    pclass = case_when(
      pclass == 1 ~ "Upper",
      pclass == 2 ~ "Middle",
      pclass == 3 ~ "Lower"
    )
  )
```

```{r}
#Change the name of columns
selected_titanic <- selected_titanic %>% 
  rename(
    Class = pclass,
    Survived = survived,
    Age = age
  )
```


### Visualizing and revealing patterns and insights   

#### Fare Categories 

```{r}
#Create a table object fare_count
fare_count <- table(selected_titanic$FareCategory)
print(fare_count)
```
**Note: the website has a tibble of 270 rows, however reports 681 in their frequency counts... 

```{r}

library(ggplot2)
# Create summary data frame of counts and percentage of
# fare category from the selected titanic
summary_fare <- selected_titanic %>%
  group_by(FareCategory)%>%
  summarize(n = n())%>%
  mutate(Fpercentage = n/sum(n) * 100)

# Create pie chart using the summary data frame
ggplot(summary_fare, aes(x = " ", y = n, fill = FareCategory))+
  geom_bar(stat = "identity", width = 1, color = "white")+
  coord_polar(theta = "y")+
  theme_void()+
  labs("Fare Distribution in Titanic Dataset", fill = "Fare Category")+
  geom_text(aes(label = paste0(round(Fpercentage, 2), "%")),
                   position = position_stack(vjust = 0.5), color = "white")
```
** Note: not particularly fond of the pie charts

#### Looking at gender distribution

```{r}
gender_count <- table(selected_titanic$sex)
print(gender_count)
```
```{r}
summary_gender <- selected_titanic %>%
  group_by(sex) %>%
  summarize(n = n()) %>%
  mutate(Gpercentage = n / sum(n) * 100)

ggplot(summary_gender, aes(x = "", y = n, fill = sex)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Gender Distribution in Titanic Dataset", fill = "Gender") +
  geom_text(aes(label = paste0(round(Gpercentage, 2), "%")),
            position = position_stack(vjust = 0.5), color = "white")
```


#### Age distribution 

```{r}
# Create a table of counts object for the Age column from the selected_titanic
# data frame
age_count <- table(selected_titanic$Age)

# Visualize the Age Distribution using pie function instead of ggplot2 library
pie(age_count, main = "Age Distribution in Titanic Dataset", 
    labels = paste(round(age_count/sum(age_count) * 100, 2), "%", sep = ""), 
    col = rainbow(length(age_count)))

# Add legend with age groups
legend("topright", names(age_count), cex = 0.8, 
       fill = rainbow(length(age_count)))
```

#### Port of embarkation distribution

```{r}
# Create a bar chart visualizing the Embarked column using functions from 
# the ggplot2 library

ggplot(selected_titanic, aes(x = embarked, fill = embarked)) +
  geom_bar() +
  theme_classic() +
  labs(x = "Port of Embarkation", y = "Number of Passengers",
       title = "Port of Embarkation Distribution in Titanic Dataset")

```

#### Ticket class distribution

```{r}
# Create a table of counts object for the Class column 
class_count <- table(selected_titanic$Class)

# Visualize the Age Distribution using pie function instead of ggplot2 library
 barplot(class_count,
    main = "Passengers' Ticket Class Distribution in Titanic Dataset", 
    xlab = "Ticket Class", 
    ylab = "Number of Passengers", 
    col = "skyblue")
```

#### Destination distribution 

```{r}
summary_destination <- selected_titanic %>%
    group_by(home.dest) %>%
    summarise(count = n()) %>%
    arrange(desc(count)) %>%
    head(10)

# Convert home.dest to a factor for better plotting
summary_destination$home.dest <- factor(summary_destination$home.dest, levels = summary_destination$home.dest)

# Create a heatmap
ggplot(summary_destination, aes(x = home.dest, y = count, fill = count)) +
  geom_tile() +
  labs(title = "Top 10 Destination Distribution of Passengers in Titanic Dataset", 
       x = "Destination", y = "Number of Passengers") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

#### Survival rate distribution

```{r}

# Create a summary data frame object for the Survived column
summary_survived <- selected_titanic %>%
  group_by(Survived) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(Spercentage = n / sum(n) * 100)

# Create a pie chart using ggplot2 library
ggplot(summary_survived, aes(x = "", y = n, fill = factor(Survived))) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Survival Distribution in Titanic Dataset", fill = "Survived") +
  geom_text(aes(label = paste0(round(Spercentage, 2), "%")), 
            position = position_stack(vjust = 0.5), color = "white") +
  scale_fill_manual(values = c("0" = "red", "1" = "darkgreen"), 
                    labels = c("Did Not Survive", "Survived"))

```

### Univariate Analyses 

#### Survival rate by category 

```{r}
# Bin the fare values into categories
selected_titanic <- titanic %>%
  mutate(FareCategory = cut(fare, 
                             breaks = c(-1, 10, 30, 100, Inf), 
                             labels = c("Low", "Medium", "High", "Very High")))

# Create a summary dataframe object
summary_sf <- selected_titanic %>%
  group_by(FareCategory, survived) %>%
  summarise(survival_rate = n(), .groups = "drop") %>%
  mutate(sfPercentage = survival_rate / sum(survival_rate) * 100)

# Create the plot using the summary data frame
ggplot(summary_sf, aes(x = FareCategory, y = sfPercentage, fill = factor(survived))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Survival Rate by Fare Category in Titanic Dataset", 
       x = "Fare Category", y = "Survival Rate (%)") +
  scale_fill_manual(values = c("0" = "red", "1" = "darkgreen"), 
                    labels = c("Did Not Survive", "Survived")) +
  theme_minimal()

```

```{r}

# Create a summary dataframe object
summary_sg <- selected_titanic %>%
  group_by(sex, survived) %>%
  summarise(count = n(), .groups = "drop") %>%  # Fixing the grouping issue
  mutate(sgPercentage = count / sum(count) * 100)  # Ensuring percentages are correctly calculated

# Create the plot
ggplot(summary_sg, aes(x = sex, y = sgPercentage, fill = factor(survived))) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Survival Rate by Gender in Titanic Dataset", 
       x = "Gender", y = "Survival Rate (%)") +
  theme_classic() +
  scale_fill_manual(values = c("0" = "red", "1" = "darkgreen"), 
                    labels = c("Did Not Survive", "Survived"))

```

#### Survival rate by age

```{r}
# Bin the Age variable into categories
selected_titanic <- selected_titanic %>%
  mutate(age_group = cut(age, 
                         breaks = c(-Inf, 18, 35, 60, Inf), 
                         labels = c("Child", "Young Adult", "Adult", "Elderly")))

# Create a summary dataframe object
summary_sa <- selected_titanic %>%
  group_by(age_group, survived) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(saPercentage = count / sum(count) * 100)

# Create the plot
ggplot(summary_sa, aes(x = age_group, y = saPercentage, fill = factor(survived))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Survival Rate by Age in Titanic Dataset",
       x = "Age Group", y = "Survival Rate (%)", fill = "Survived") +
  theme_classic() +
  scale_fill_manual(values = c("0" = "red", "1" = "darkgreen"),
                    labels = c("Did Not Survive", "Survived"))

```

#### Survival rate by class 
```{r}
 summary_sc <- selected_titanic %>%
  group_by (pclass, survived) %>%
  summarise(count = n())%>%
  mutate(scPercentage = count/sum(count) * 100)

ggplot(summary_sc, aes (x = pclass, y = scPercentage, fill = survived)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Survival Rate by Ticket Class in Titanic Dataset",
           x = "Ticket Class", y = "Survival Rate (%)")+
  theme_minimal()
```

#### Survival rate by deportation port 

```{r}
# Create the summary dataframe
summary_sp <- selected_titanic %>%
  group_by(embarked, survived) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(spPercentage = count / sum(count) * 100)

# Create the plot
ggplot(summary_sp, aes(x = embarked, y = spPercentage, fill = factor(survived))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Survival Rate by Port of Embarkation in Titanic Dataset", 
       x = "Port of Embarkation", y = "Survival Rate (%)") +
  theme_classic() +
  scale_fill_manual(values = c("0" = "red", "1" = "darkgreen"), 
                    labels = c("Did Not Survive", "Survived"))

```

### Commentary 

Overall, I enjoyed working through this lab as I found the topic to peak my interest. However, for the purposes of our class I believe that there are quite a few changes that could be made. 

1) There was a lot of debugging that I had to do with the code provided. Much of this was for minor things such as sloppy capitalization, functions not working, etc. However, this could be very frustrating especially if it is used as an early lab and students are not well versed in debugging quite yet. 

2) There was a lot of text that needed to be paraphrased/cut out. 

3) This lab seemed to go on forever, and it would have benefited from clear/delineated steps. 

4) There was an over reliance on pie charts, which as we have learned are not the best for presenting data. 

5) It was oddly difficult to find the dataset that was being used - no link given, vague instructions... took much longer than needed. 

6) Better context could have been required for the steps, also there were some functions that were just given without proper activity.

7) Lab could have been a bit more hands on, I genuinely felt like I was just copy and pasting code/didn't write a single thing. 

8) There appear to be some errors with the number of people left in the dataset/something isn't adding up see notes on counts above.  


For the next portfolio piece, my aim is to create a lab that would fit into the visualization portion of our class more fluidly, and improve some the areas above. 


:) 





