---
title: "Portfolio 3 - Fueling My Titanic Obsession"
---

> The goal of this project is to evaluate and make improvements to a Titanic Lab found online. https://medium.com/@datathon/r-for-data-science-data-wrangling-d125285a6876 

### Eliminating early "fluff"
> Much of the current lab has information that is unnecessary for where students are at this point in the semester. For example, the site includes installing R directions and explains the layout of R. For a visualization lab for this class, these instructions are unnessecary.

### Importing Data

> The clarity of this step could be improved. Here are some new steps:

1) Go to Kaggle Titanic Dataset - there are a lot of datasets on titanic out there, however the lab was not clear which was being used... so i believe this is the best match 
https://www.kaggle.com/datasets/ibooth1/titanic3 

2) Download the dataset in the format you would like, I did excel csv

3) Extract the data set to your portfolio folder

4) Install/load packages 

```{r}
library(tidyr)
library(readxl)
library(Rcpp) # needed to load Amelia
library(Amelia) # package for missing data/imputation
library(dplyr)
```

5) Read in data 
```{r}
library(readxl)
titanic <- read_xls("titanic3.xls")

```

### Exploring and cleaning dataset 

```{r}
str(titanic)
```
```{r}
as_tibble(titanic)
```


#### Checking for missing values 
```{r}
missmap(titanic, col = c("red", "green"))
```
21% of data is missing 

```{r}
#Select relevant columns for the analysis
selected_titanic <- titanic %>%
  select(age, pclass, sex, survived, embarked, fare, cabin, parch, sibsp)
```

Reassess missing map
```{r}
missmap(selected_titanic, col = c("red", "green"))
```
Missing is now 11% 

```{r}
#Drop all the missing values
selected_titanic <- drop_na(selected_titanic)
```
```{r}
missmap(selected_titanic, col = c("red", "green"))
```

```{r}
as_tibble(selected_titanic)
```


```{r}
#merge columns parch and sibsp to create a new column, FamilySize
selected_titanic$FamilySize <- selected_titanic$sibsp
 + selected_titanic$parch + 1

#Categorize the fare column and assign label to each category
selected_titanic$FareCategory <- cut(selected_titanic$fare, breaks = c(0, 10, 20, 50, 100, Inf), labels = c("Lowest", "Lower Middle", "Upper Middle", "Higher", "Highest"))
```

```{r}
View(selected_titanic)
```

```{r}
#Remove irrelevant columns
selected_titanic <- selected_titanic %>% 
  select(-fare, -parch, -sibsp)
```

```{r}
#Change the values of columns survived, age, and pclass

selected_titanic <- selected_titanic %>%
  mutate(
    survived = ifelse(survived == 0, "No", "Yes"),

    age = ifelse(age >= 18, "Adult", "Child"),

    pclass = case_when(
      pclass == 1 ~ "Upper",
      pclass == 2 ~ "Middle",
      pclass == 3 ~ "Lower"
    )
  )
```

```{r}
#Change the name of columns
selected_titanic <- selected_titanic %>% 
  rename(
    Class = pclass,
    Survived = survived,
    Age = age
  )
```


### Visualizing and revealing patterns and insights   

#### Fare Categories 

```{r}
#Create a table object fare_count
fare_count <- table(selected_titanic$FareCategory)
print(fare_count)
```
**Note: the website has a tibble of 270 rows, however reports 681 in their frequency counts... 

```{r}

library(ggplot2)
# Create summary data frame of counts and percentage of
# fare category from the selected titanic
summary_fare <- selected_titanic %>%
  group_by(FareCategory)%>%
  summarize(n = n())%>%
  mutate(Fpercentage = n/sum(n) * 100)

# Create pie chart using the summary data frame
ggplot(summary_fare, aes(x = " ", y = n, fill = FareCategory))+
  geom_bar(stat = "identity", width = 1, color = "white")+
  coord_polar(theta = "y")+
  theme_void()+
  labs("Fare Distribution in Titanic Dataset", fill = "Fare Category")+
  geom_text(aes(label = paste0(round(Fpercentage, 2), "%")),
                   position = position_stack(vjust = 0.5), color = "white")
```
** Note: not particularly fond of the pie charts

#### Looking at gender distribution

```{r}
gender_count <- table(selected_titanic$sex)
print(gender_count)
```
























