---
title: "Portfolio 3 - Feeding my Titanic Obsession"
---

> The goal of this project is to evaluate and make improvements to a Titanic Lab found online. https://medium.com/@datathon/r-for-data-science-data-wrangling-d125285a6876 

### Eliminating early "fluff"
> Much of the current lab has information that is unnecessary for where students are at this point in the semester. For example, the site includes installing R directions and explains the layout of R. For a visualization lab for this class, these instructions are unnessecary.

### Importing Data

> The clarity of this step could be improved. Here are some new steps:

1) Go to Kaggle Titanic Dataset 
https://www.kaggle.com/datasets/yasserh/titanic-dataset?resource=download 

2) Download the dataset in the format you would like, I did excel csv

3) Extract the data set to your portfolio folder

4) Install/load packages 

```{r}
library(tidyr)
library(readxl)
library(Rcpp) # needed to load Amelia
library(Amelia) # package for missing data/imputation
```

5) Read in data 
```{r}
titanic <- read.csv("Titanic-Dataset.csv")
```


### Exploring and cleaning dataset 

