---
title: "Portfolio 4 - Lying Graph for Thesis "
---

The goal of this project is to load my SPSS data into R and create a box and whiskers plot representing lying within each scenario. 

##Step 1: Reading in SPSS Data 

```{r}
library(haven)
library(tidyr)
library(tidyverse)
```


```{r}
Thesis <- read_spss("C:/Users/crawl/OneDrive/Documents/GitHub/Portfolio/ThesisAllDataNew.sav")
```

##Step 2: Creating Plots per scenario 
> Goal here is to have a box and wiskers plot for each scenario (Romantic Partner-RP and Twenty Dollars-TD) with the dots visible for each data point. 

```{r}
ggplot(Thesis, aes(x = "", y = RPLieAvg)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray")+
  geom_jitter(width = 0.1, size = 2, alpha = 0.6, color = "Black") +
  labs(x = NULL, y = "RPLieAvg") +
  theme_minimal()
```


```{r}
ggplot(Thesis, aes(x = "", y = TDLieAvg)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray")+
  geom_jitter(width = 0.1, size = 2, alpha = 0.6, color = "Black") +
  labs(x = NULL, y = "TDLieAvg") +
  theme_minimal()
```

## Step 3: Combining the scenarios into one plot


```{r}
library(tidyr)
library(dplyr)


Thesis_long <- Thesis %>%
  pivot_longer(cols = c(RPLieAvg, TDLieAvg),
               names_to = "Scenario",
               values_to = "LieAvg")

```

```{r}

library(extrafont)

# Run once to import system fonts (takes a bit)
font_import(prompt = FALSE)

# Register fonts
loadfonts(device = "win")  # or use `device = "pdf"` or "postscript" depending on output

```


```{r}
library(ggplot2)
library(extrafont)

ggplot(Thesis_long, aes(x = Scenario, y = LieAvg)) +
  geom_boxplot(outlier.shape = NA, fill = "lightgray") +
  geom_jitter(width = 0.2, size = 2, alpha = 0.6, color = "black") +
  labs(x = "Scenario", y = "Average Lying Score") +
  theme_minimal(base_family = "Times New Roman")
```



