---
title: "Portfolio 5 - Pixar Films"
---

> The goal here is to explore the Pixar Films data from TidyTuesday and see if I can find anything of interest. 

```{r}
##install.packages("pixarfilms")
```

```{r}
library("pixarfilms")
library(tidyr)
library(dplyr)

```

```{r}
pixar <- pixar_films
```

## Runtime Trends 

```{r}
mean(pixar$run_time, na.rm = TRUE)
```
The average runtime for a Pixar film is around 105 minutes - how has that changed over time? 

```{r}
library(ggplot2)
```

```{r}
pixar %>%  
  ggplot(aes(x = release_date, y = run_time)) +
  geom_line() +
  geom_point() +
  labs(title = "Pixar Film Runtimes Over Time", x = "Year", y = "Run Time ")
```

It appears that there was a large spike shortly after 2021 - I am curious to know which movie had such a long run time. 

```{r}
ggplot(pixar, aes(x = reorder(film, run_time), y = run_time)) +
  geom_col() +
  coord_flip() +
  labs(title = "Pixar Film Runtimes", x = "Film", y = "Run Time")
```
```{r}
pixar[which.max(pixar$run_time), ]
pixar[which.min(pixar$run_time), ]

```


Apparently Luca has an extremely long run time, compared to the other Pixar films... perhaps I have to watch it now...

Toy Story has the shortest run time - remains superior though...

## Rating Trends 

Let's see how run time varies by rating 

```{r}
pixar %>%
  group_by(film_rating) %>%
  summarise(mean_runtime = mean(run_time), count = n())
```
The majority of the movies are G rated and have an average run time of around 100 mins followed by PG rated moves with a similar run time (avg of 101). Not extremely surprised by these results. Let's plot this ...

```{r}
pixar %>%  
  ggplot(aes(x = film_rating, fill = film_rating)) +
  geom_bar() +
  labs(title = "Distribution of Film Ratings", x = "Rating", y = "Count")
```

I am curious to know if ratings have become "less restrictive" or rather if more movies have been made recently that have more "severe" ratings

```{r}
pixar %>%  
  ggplot(aes(x = release_date, y = film_rating, color = film_rating)) +
  geom_point() +
  labs(title = "Pixar Film Runtimes Over Time", x = "Release Date", y = "Film Rating")
```
 
Not particularly informative, but it does look like most early films were G rated where as around 2005 it looks like PG movies were introduced.
 
## Years Since Release 

```{r}
library(lubridate)
```

```{r}
pixarnew <- pixar %>%
  mutate(years_since_release = year(Sys.Date()) -  year(release_date))
```

```{r}
pixarnew %>%
  group_by(film_rating) %>%
  summarise(mean_years_old = mean(years_since_release, na.rm = TRUE))
```
The G rated films on average tend to be older than the PG rated films. 

Let's see if films are getting shorter or longer...  

```{r}
ggplot(pixarnew, aes(x = years_since_release, y = run_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Do Pixar Films Get Longer Over Time?", y = "RunTime", x = "Years Since Release")

```
It looks like no - Pixar films have not gotten longer over time, rather more recent movies appear to be longer, but I am curious to see this without the two "outliers"


```{r}
pixarfilter <- pixarnew %>%  
  filter(run_time < 130)
```

```{r}
ggplot(pixarfilter, aes(x = years_since_release, y = run_time)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Do Pixar Films Get Longer Over Time?", y = "RunTime", x = "Years Since Release")

```

Still appears to be a negative relationship! 

Overall, I think this was a fun little experiment. However, I was hoping that there was going to be more to the data like info about characters(male/female/other? animal/human? princess? etc.), themes of plots, viewer ratings, directors, or funding. Nevertheless, explored the data and made some cool graphs! 
