---
title: "Portfolio 5"
---

> The goal here is to explore the Pixar Films data from TidyTuesday and see if I can find anything of interest. 

```{r}
##install.packages("pixarfilms")
```

```{r}
library("pixarfilms")
library(tidyr)
library(dplyr)

```

```{r}
pixar <- pixar_films
```

## Runtime Trends 

```{r}
mean(pixar$run_time, na.rm = TRUE)
```
The average runtime for a Pixar film is around 105 minutes - how has that changed over time? 

```{r}
library(ggplot2)
```

```{r}
pixar %>%  
  ggplot(aes(x = release_date, y = run_time)) +
  geom_line() +
  geom_point() +
  labs(title = "Pixar Film Runtimes Over Time", x = "Year", y = "Run Time ")
```

It appears that there was a large spike shortly after 2021 - I am curious to know which movie had such a long run time. 

```{r}
ggplot(pixar, aes(x = reorder(film, run_time), y = run_time)) +
  geom_col() +
  coord_flip() +
  labs(title = "Pixar Film Runtimes", x = "Film", y = "Run Time")
```
```{r}
pixar[which.max(pixar$run_time), ]
pixar[which.min(pixar$run_time), ]

```


Apparently Luca has an extremely long run time, compared to the other Pixar films... perhaps I have to watch it now...

Toy Story has the shortest run time - remains superior though...

## Rating Trends 

Let's see how run time varies by rating 

```{r}
pixar %>%
  group_by(film_rating) %>%
  summarise(mean_runtime = mean(run_time), count = n())
```
The majority of the movies are G rated and have an average run time of around 100 mins followed by PG rated moves with a similar run time (avg of 101). Not extremely surprised by these results. Let's plot this ...

```{r}
pixar %>%  
  ggplot(aes(x = film_rating, fill = film_rating)) +
  geom_bar() +
  labs(title = "Distribution of Film Ratings", x = "Rating", y = "Count")
```

I am curious to know if ratings have become "less restrictive" or rather if more movies have been made recently that have more "severe" ratings

```{r}
pixar %>%  
  ggplot(aes(x = release_date, y = film_rating, color = film_rating)) +
  geom_point() +
  labs(title = "Pixar Film Runtimes Over Time", x = "Release Date", y = "Film Rating")
```
 
Not particularly informative, but it does look like most early films were G rated where as around 2005 it looks like PG movies were introduced.
 
## Years Since Release 

```{r}
pixar %>% 
  mutate()
```

