---
title: "Portfolio 7"
---

> The goal of this project is to explore the Great American Coffee Tasting data set from Tidy Tuesday.

```{r}
library(tidytuesdayR)
```

```{r}
tuesdata <- tidytuesdayR::tt_load(2024, week = 20)

coffee_survey <- tuesdata$coffee_survey
```
## Coffee Drinker's Demographics

```{r}
library(tidyverse)
library(cluster)
library(dplyr)
library(ggplot2)
```

Let's see if coffee drinking habits vary by a variety of factors including: gender, employment status, education, number of children, political affiliation, and ethnicity. 

### Gender

```{r}
ggplot(coffee_survey, aes(x = cups, fill = gender)) +
  geom_bar() +
  labs(title = "Coffee Consumption by Gender",
       x = "Cups per Day", y = "Count", fill = "Gender") +
  theme_minimal()

```

```{r}
ggplot(coffee_survey, aes(x = cups, fill = gender)) +
  geom_bar(position = "fill") +  # stack bars by proportion
  labs(title = "Coffee Consumption by Gender (Proportions)",
       x = "Cups per Day", y = "Proportion") +
  theme_minimal()


```
```{r}
ggplot(coffee_survey, aes(x = cups)) +
  geom_bar(fill = "steelblue") +
  facet_wrap(~ gender) +
  labs(title = "Cups of Coffee per Day by Gender",
       x = "Cups per Day", y = "Count") +
  theme_minimal()

```

```{r}
coffee_survey %>% 
  count(gender, cups) %>% 
  arrange(cups, gender)
```


```{r}
bycups <- coffee_survey %>%
  mutate(
    cups_numeric = case_when(
      cups == "Less than 1" ~ 0.5,
      cups == "More than 4" ~ 5,
      str_detect(cups, "^[0-9]+$") ~ as.numeric(cups),
      TRUE ~ NA_real_
    )
  )
```

```{r}
bycups %>%
  group_by(gender) %>%
  summarise(
    average_cups = mean(cups_numeric, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(average_cups))

```

It appears that a greater number of males tend to drink coffee than any other gender category. However, those who indicated "Other" as their gender tended to drink a greater number of cups than the other groups. 

## Employment and Education 

```{r}
bycups %>%
  group_by(employment_status) %>%
  summarise(
    average_cups = mean(cups_numeric, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(average_cups))

```

Those who are retired drink the greatest number of cups of coffee - which surprised me since I associate coffee with work and productivity. However, those who are employed full time make up the largest portion of the sample and drink roughly 1.82 cups of coffee a day on average. 

```{r}
bycups %>%  
  group_by(employment_status) %>%
  summarise(average_cups = mean(cups_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(employment_status, -average_cups), y = average_cups)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Average Cups per Day by Employment Status",
       x = "Employment Status", y = "Average Cups per Day") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
bycups %>% 
  group_by(education_level) %>%
  summarise(
    average_cups = mean(cups_numeric, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(average_cups))

```
The majority of individuals have a Bachelor's degree - drinking on average 1.72 cups a coffee a day. Those with Doctorate or professional degrees drink the most coffee on average at 1.93 cups a day. 

```{r}
bycups %>%  
    group_by(education_level) %>%
  summarise(average_cups = mean(cups_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(education_level, -average_cups), y = average_cups)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Average Cups per Day by Education Level",
       x = "Education Level", y = "Average Cups per Day") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Number of Children 

```{r}
bycups %>%
  group_by(number_children) %>%
  summarise(
    average_cups = mean(cups_numeric, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(number_children)

```
```{r}
bycups %>%
  group_by(number_children) %>%
  summarise(average_cups = mean(cups_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(number_children), y = average_cups)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Average Cups per Day by Number of Children",
       x = "Number of Children", y = "Average Cups per Day") +
  theme_minimal()

```

Looks like those who have 3 or more children are more likely to consume more coffee on average, as compared to those with fewer children. 


## Political Affiliation 

```{r}
bycups %>%
  group_by(political_affiliation) %>%
  summarise(
    average_cups = mean(cups_numeric, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(average_cups))

```

```{r}
bycups %>%
  group_by(political_affiliation) %>%
  summarise(average_cups = mean(cups_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(political_affiliation, -average_cups), y = average_cups)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Average Cups per Day by Political Affiliation",
       x = "Political Affiliation", y = "Average Cups per Day") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

It looks like Republicans and Independents drink roughly the same amount of coffee. Democrats tend to drink slightly less coffee. 

## Ethnicity 
```{r}
bycups %>%
  group_by(ethnicity_race) %>%
  summarise(
    average_cups = mean(cups_numeric, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(average_cups))

```
```{r}
bycups %>%
  group_by(ethnicity_race) %>%
  summarise(average_cups = mean(cups_numeric, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(ethnicity_race, -average_cups), y = average_cups)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Average Cups per Day by Ethnicity",
       x = "Ethnicity", y = "Average Cups per Day") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
Native American/Alaska Native individuals tend to drink the most cups of coffee on average! 

## Preferred Coffee and Coffee Profiles 

```{r}
coffee_survey %>%
  count(prefer_overall) %>%
  ggplot(aes(x = prefer_overall, y = n, fill = prefer_overall)) +
  geom_bar(stat = "identity") +
  labs(title = "Most Preferred Coffee (Overall Taste)",
       x = "Coffee Sample", y = "Number of People") +
  theme_minimal() +
  theme(legend.position = "none")

```
The sample far preferred Coffee D - now I want to try the coffees...


```{r}
# Prepare data
coffee_long <- coffee_survey %>%
  summarise(
    A_bit = mean(coffee_a_bitterness, na.rm = TRUE),
    A_acid = mean(coffee_a_acidity, na.rm = TRUE),
    A_pref = mean(coffee_a_personal_preference, na.rm = TRUE),
    B_bit = mean(coffee_b_bitterness, na.rm = TRUE),
    B_acid = mean(coffee_b_acidity, na.rm = TRUE),
    B_pref = mean(coffee_b_personal_preference, na.rm = TRUE),
    C_bit = mean(coffee_c_bitterness, na.rm = TRUE),
    C_acid = mean(coffee_c_acidity, na.rm = TRUE),
    C_pref = mean(coffee_c_personal_preference, na.rm = TRUE),
    D_bit = mean(coffee_d_bitterness, na.rm = TRUE),
    D_acid = mean(coffee_d_acidity, na.rm = TRUE),
    D_pref = mean(coffee_d_personal_preference, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = everything(), names_to = "coffee_attr", values_to = "rating") %>%
  separate(coffee_attr, into = c("coffee", "attribute"), sep = "_")

```


```{r}
ggplot(coffee_long, aes(x = attribute, y = rating, fill = coffee)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Ratings for Each Coffee Sample",
       x = "Attribute", y = "Average Rating") +
  theme_minimal()
```
Coffee D - the preferred coffee - is the most acidic and not particularly bitter. Coffee D and A have similar ratings although as seen above, more people preferred coffee D.



```{r}
ggplot(coffee_long, aes(x = attribute, y = rating, fill = attribute)) +
  geom_col() +
  facet_wrap(~ coffee) +
  labs(title = "Attribute Ratings per Coffee Sample",
       x = "Attribute", y = "Average Rating") +
  theme_minimal()

```

Here is a profile of each coffee: 
A and D are similar in Acidity and Bitterness 
C and B are similar in Acidity and Bitterness 


## Predicting the Perfect Cup


```{r}

coffeepred <- multinom(prefer_overall ~
                          coffee_a_bitterness + coffee_a_acidity + coffee_a_personal_preference +
                          coffee_b_bitterness + coffee_b_acidity + coffee_b_personal_preference +
                          coffee_c_bitterness + coffee_c_acidity + coffee_c_personal_preference +
                          coffee_d_bitterness + coffee_d_acidity + coffee_d_personal_preference +
                          strength + brew + age,
                        data = df_model)

summary(multi_model)
```





