---
title: "Portfolio 8 - Thesis EVT Correlations"
---

> The goal of this project is to run and graph the Expectancy-Value theory correalations for my thesis. 

```{r}
library(haven)
library(tidyr)
library(tidyverse)
```

"C:\Users\crawl\OneDrive - Wake Forest University\GitHub\Portfolio\ThesisAllDataNewNew.sav"

```{r}
Thesis <- read_spss("C:/Users/crawl/OneDrive - Wake Forest University/GitHub/Portfolio/ThesisAllDataNewNew.sav")
```

### Hypothesis 1 
> Lying, AV, OV 

```{r}
  cor.test(Thesis$RPLieAvg, Thesis$RP_AV, method = "pearson")
```
```{r}
library(ggplot)
```

```{r}
ggplot(Thesis, aes(x = RPLieAvg, y = RP_AV)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "RP Lying Average (RPLieAvg)",
    y = "RP Action Value (RPAV)",
    title = "Correlation between RPLieAverage and AV"
  ) +
  theme_minimal()
```


```{r}
  cor.test(Thesis$RPLieAvg, Thesis$RP_OV, method = "pearson")
```

```{r}
ggplot(Thesis, aes(x = RPLieAvg, y = RP_OV)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "RP Lying Average (RPLieAvg)",
    y = "RP Outcome Value (RPOV)",
    title = "Correlation between RPLieAverage and OV"
  ) +
  theme_minimal()
```




```{r}
  cor.test(Thesis$TDLieAvg, Thesis$TD_AV, method = "pearson")
```

```{r}
ggplot(Thesis, aes(x = TDLieAvg, y = TD_AV)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "TD Lying Average (TDLieAvg)",
    y = "TD Action Value (TDAV)",
    title = "Correlation between TDLieAverage and AV"
  ) +
  theme_minimal()
```

```{r}
  cor.test(Thesis$TDLieAvg, Thesis$TD_OV, method = "pearson")
```

```{r}
ggplot(Thesis, aes(x = TDLieAvg, y = TD_OV)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "TD Lying Average (TDLieAvg)",
    y = "TD Outcome Value (TDOV)",
    title = "Correlation between TDLieAverage and OV"
  ) +
  theme_minimal()
```

To the degree that an individual values the act of attempting to deceive others and successfully deceiving others they are more inclined to lie. 

### Hypothesis 2 
> Lying and EBL 


```{r}
  cor.test(Thesis$RPLieAvg, Thesis$RP_PLE_FC, method = "pearson")
```

```{r}
ggplot(Thesis, aes(x = RPLieAvg, y = RP_PLE_FC)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "RP Lying Average (RPLieAvg)",
    y = "RP EBL",
    title = "Correlation between RPLieAverage and the Expected Benefit of Lying"
  ) +
  theme_minimal()
```


```{r}
  cor.test(Thesis$TDLieAvg, Thesis$TD_PLE_FC, method = "pearson")
```

```{r}
ggplot(Thesis, aes(x = TDLieAvg, y = TD_PLE_FC)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "TD Lying Average (TDLieAvg)",
    y = "TD PLE",
    title = "Correlation between TDLieAverage and the Expected Benefit of Lying"
  ) +
  theme_minimal()
```

Individuals are more likely to lie in a given situation if they believe that their most valued goals will be accomplished via deception. 

### Hypothesis 3 

```{r}
model1_step1 <- lm(RPLieAvg ~ RP_PLE_FC + RP_AOE_C, data = Thesis)
summary(model1_step1)

model1_step2 <- lm(RPLieAvg ~ RP_PLE_FC + RP_AOE_C + RP_Int, data = Thesis)
summary(model1_step2)
```

```{r}
model2_step1 <- lm(TDLieAvg ~ TD_PLE_FC + TD_AOE_C, data = Thesis)
summary(model2_step1)


model2_step2 <- lm(TDLieAvg ~ TD_PLE_FC + TD_AOE_C + TD_Int, data = Thesis)
summary(model2_step2)
```

Sig positive slope for interactions in each scenario - individual with a high AOE is significantly more likely to lie, compared to an individual with a low AOE, because they perceive there to be a stronger connection between the expected benefit of lying (EBL) and deception itself. 
