---
title: "Portfolio 8 - Thesis EVT Correlations"
---

> The goal of this project is to run and graph the Expectancy-Value theory correalations for my thesis. 

```{r}
library(haven)
library(tidyr)
library(tidyverse)
```



```{r}
Thesis <- read_spss("C:/Users/crawl/OneDrive/Documents/GitHub/Portfolio/ThesisAllDataNew.sav")
```


### Hypothesis 1 
> Lying, AV, OV 

```{r}
  cor.test(Thesis$RPLieAvg, Thesis$RP_AV, method = "pearson")
```
```{r}
library(ggplot)
```

```{r}
ggplot(Thesis, aes(x = RPLieAvg, y = RP_AV)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "RP Lying Average (RPLieAvg)",
    y = "RP Action Value (RPAV)",
    title = "Correlation between RPLieAverage and AV"
  ) +
  theme_minimal()
```


```{r}
  cor.test(Thesis$RPLieAvg, Thesis$RP_OV, method = "pearson")
```

```{r}
ggplot(Thesis, aes(x = RPLieAvg, y = RP_OV)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "RP Lying Average (RPLieAvg)",
    y = "RP Outcome Value (RPOV)",
    title = "Correlation between RPLieAverage and OV"
  ) +
  theme_minimal()
```




```{r}
  cor.test(Thesis$TDLieAvg, Thesis$TD_AV, method = "pearson")
```

```{r}
ggplot(Thesis, aes(x = TDLieAvg, y = TD_AV)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "TD Lying Average (TDLieAvg)",
    y = "TD Action Value (TDAV)",
    title = "Correlation between TDLieAverage and AV"
  ) +
  theme_minimal()
```

```{r}
  cor.test(Thesis$TDLieAvg, Thesis$TD_OV, method = "pearson")
```

```{r}
ggplot(Thesis, aes(x = TDLieAvg, y = TD_OV)) +
  geom_jitter(alpha = 0.3) + 
  geom_smooth(method = "lm", se = TRUE) +  
  labs(
    x = "TD Lying Average (TDLieAvg)",
    y = "TD Outcome Value (TDOV)",
    title = "Correlation between TDLieAverage and OV"
  ) +
  theme_minimal()
```

To the degree that an individual values the act of attempting to deceive others and successfully deceiving others they are more inclined to lie. 

### Hypothesis 2 
> Lying and EBL 


Individuals are more likely to lie in a given situation if they believe that their most valued goals will be accomplished via deception. 

### Hypothesis 3 
